---
title: "EmpAttrition"
output: html_document
---

### Load the dataset
```{r}
data_dir <- "~/Desktop/Springboard/EmployeeAttrition"
term_data <- read.csv(file.path(data_dir,"TermData_org.csv"),header = TRUE)
MYdataset <- term_data
str(MYdataset)
library(plyr)
library(dplyr)
 
knitr::opts_chunk$set(echo = TRUE)
```

### Look at the data quality

```{r}
summary(MYdataset)
```


### Termination Proportion

```{r}
StatusCount<- as.data.frame.matrix(MYdataset %>%
group_by(STATUS_YEAR) %>%
select(STATUS) %>%
table())
StatusCount$TOTAL<-StatusCount$ACTIVE + StatusCount$TERMINATED
StatusCount$PercentTerminated <-StatusCount$TERMINATED/(StatusCount$TOTAL)*100
StatusCount
mean(StatusCount$PercentTerminated)

```


###  Plots by Business Unit


```{r , echo=FALSE}
library(ggplot2)
ggplot() + geom_bar(aes(y = ..count..,x =as.factor(BUSINESS_UNIT),fill = as.factor(STATUS)),data=MYdataset,position = position_stack())

```

###  Plots by Term Type

```{r , echo=FALSE}

TerminatesData<- as.data.frame(MYdataset %>%
filter(STATUS=="TERMINATED"))
ggplot() + geom_bar(aes(y = ..count..,x =as.factor(STATUS_YEAR),fill = as.factor(termtype_desc)),data=TerminatesData,position = position_stack())
```


###  Build the model

#### Partition the data

```{r }

library(magrittr)
building <- TRUE
 scoring  <- ! building
 set.seed(42)
 MYnobs <- nrow(MYdataset)
 MYsample <- MYtrain <- subset(MYdataset,STATUS_YEAR<=2014)
 MYvalidate <- NULL
 MYtest <- subset(MYdataset,STATUS_YEAR== 2015)
 
 # The following variable selections have been noted.
 MYinput <- c("age", "length_of_service",    "gender_full", "STATUS_YEAR", "BUSINESS_UNIT")
 MYnumeric <- c("age", "length_of_service", "STATUS_YEAR")

MYcategoric <- c("gender_full", "BUSINESS_UNIT")
MYtarget  <- "STATUS"
MYrisk   <-  NULL
MYident <-  "EmployeeID"
MYignore <- c("recorddate_key", "birthdate_key", "orighiredate_key", "terminationdate_key", "city_name", "gender_short", "termreason_desc", "termtype_desc","department_name","job_title", "store_name")
MYweights <-  NULL
MYTrainingData<-MYtrain[c(MYinput, MYtarget)]
MYTestingData<-MYtest[c(MYinput, MYtarget)]
```

###  Model Selection  

#### Decision Tree
```{r}
#install.packages('rattle')
#install.packages('rpart')
#install.packages('rpart.plot')
#install.packages('RColorBrewer')
#library(rattle)
library(rpart.plot)
library(RColorBrewer)
library(rpart, quietly=TRUE)
set.seed(42)

# Build the Decision Tree model.
MYrpart <- rpart(STATUS ~ .,
                   data=MYtrain[, c(MYinput, MYtarget)],
                   method="class",
                   parms=list(split="information"),
                   control=rpart.control(usesurrogate=0, 
                                         maxsurrogate=0))
#fancyRpartPlot(MYrpart, main="Decision Tree MFG10YearTerminationData $ STATUS")
```

#### Random Forest
```{r}
library(randomForest, quietly=TRUE)

set.seed(42)
MYrf <- randomForest(STATUS ~ .,
                                   data=MYtrain[c(MYinput, MYtarget)],
                                   ntree=500,
                                   mtry=2,
                                   importance=TRUE,
                                   na.action=randomForest::na.roughfix,
                                   replace=FALSE)
MYrf
# Calculate the Area Under the Curve (AUC).
#install.packages("pROC")
pROC::roc(MYrf$y, as.numeric(MYrf$predicted))

# Calculate the AUC Confidence Interval.
pROC::ci.auc(MYrf$y, as.numeric(MYrf$predicted))

# List the importance of the variables.

rn <- round(randomForest::importance(MYrf), 2)
rn[order(rn[,3], decreasing=TRUE),]
```

#### Adaoost

```{r}
set.seed(42)
#install.packages("ada")
library(ada)
MYada <- ada(STATUS ~ .,
                  data=MYtrain[c(MYinput, MYtarget)],
                  control=rpart::rpart.control(maxdepth=30,
                                               cp=0.010000,
                                               minsplit=20,
                                               xval=10),
                                               iter=50)
# Print the results of the modelling.

print(MYada)
round(MYada$model$errs[MYada$iter,], 2)
```

#### Linear Models

```{r}
# Build a Regression model.

MYglm <-glm(STATUS ~ .,
             data=MYtrain[c(MYinput, MYtarget)],
             family=binomial(link="logit"))

# Generate a textual view of the Linear model.

print(summary(MYglm))

cat(sprintf("Log likelihood: %.3f (%d df)\n",
            logLik(MYglm)[1],
            attr(logLik(MYglm), "df")))
cat(sprintf("Null/Residual deviance difference: %.3f (%d df)\n",
            MYglm$null.deviance-MYglm$deviance,
            MYglm$df.null-MYglm$df.residual))

cat(sprintf("Chi-square p-value: %.8f\n",
            dchisq(MYglm$null.deviance-MYglm$deviance,
                   MYglm$df.null-MYglm$df.residual)))
cat(sprintf("Pseudo R-Square (optimistic): %.8f\n",
            cor(MYglm$y, MYglm$fitted.values)))

cat('\n==== ANOVA ====\n\n')
print(anova(MYglm, test="Chisq"))
```

### Evaluate Models    

#### Evaluate Decision Tree Using AUC
```{r}
#install.packages("ROCR")
library(ROCR)
library(ggplot2,quietly = TRUE)

# Generate an ROC Curve for the rpart model on MFG10YearTerminationData [test].

MYpr <- predict(MYrpart, newdata=MYtest[c(MYinput, MYtarget)])[,2]

# Remove observations with missing target.

no.miss <-  na.omit(MYtest[c(MYinput, MYtarget)]$STATUS)
miss.list<- attr(no.miss, "na.action")
attributes(no.miss)<- NULL

if (length(miss.list)){pred <- prediction(MYpr[-miss.list], no.miss)} else {pred<- prediction(MYpr, no.miss)}

pe <- performance(pred, "tpr", "fpr")
au <- performance(pred, "auc")@y.values[[1]]
pd <- data.frame(fpr=unlist(pe@x.values), tpr=unlist(pe@y.values))
ggplot(pd, aes(x=fpr, y=tpr)) + geom_line(colour = "red") + labs(x= "False Positive Rate", y = "True Positive Rate")+ggtitle("ROC Curve Decision Tree MyDataset [test] STATUS") + geom_line(data=data.frame(), aes(x=c(0,1), y=c(0,1)), colour="grey") + annotate("text", x=0.50, y=0.00, hjust=0, vjust=0, size=5,label=paste("AUC =", round(au, 2)))
```

##### Evaluate Adaboost Model 
```{r}
# Calculate the area under the curve for the plot.
# Remove observations with missing target.
no.miss  <-  na.omit(MYtest[c(MYinput, MYtarget)]$STATUS)
miss.list <- attr(no.miss, "na.action")
attributes(no.miss) <- NULL

if (length(miss.list)) {pred <- prediction(MYpr[-miss.list], no.miss)} else {pred <- prediction(MYpr, no.miss)}

performance(pred,"auc")

library(ROCR)
library(ggplot2, quietly=TRUE)

# Generate an ROC Curve for the ada model on MFG10YearTerminationData [test].
MYpr <- predict(MYada, newdata=MYtest[c(MYinput, MYtarget)], type="prob")[,2]
 no.miss <-   na.omit(MYtest[c(MYinput, MYtarget)]$STATUS)
 miss.list <- attr(no.miss, "na.action")
 attributes(no.miss) <- Brown <- NULL
 if (length(miss.list)){pred <- prediction(MYpr[-miss.list], no.miss)} else { pred <- prediction(MYpr, no.miss)}
 
pe <- performance(pred, "tpr", "fpr")
au <- performance(pred, "auc")@y.values[[1]]
pd <- data.frame(fpr=unlist(pe@x.values), tpr=unlist(pe@y.values))

ggplot(pd, aes(x=fpr, y=tpr)) + geom_line(colour="red") +labs(x= "False Positive Rate", y = "True Positive Rate") +ggtitle("ROC Curve Ada Boost MyDataset [test] STATUS") +geom_line(data=data.frame(), aes(x=c(0,1), y=c(0,1)), colour="grey") + annotate("text", x=0.50, y=0.00, hjust=0, vjust=0, size=5,label=paste("AUC =", round(au, 2)))
```


##### Evaluate Random Forest Model
``` {r}
# Remove observations with missing target.
no.miss  <-  na.omit(MYtest[c(MYinput, MYtarget)]$STATUS)
miss.list <- attr(no.miss, "na.action")
attributes(no.miss) <-Brown <- NULL
if (length(miss.list)) {pred <- prediction(MYpr[-miss.list], no.miss)} else {pred <- prediction(MYpr, no.miss)}
performance(pred, "auc")

library(ROCR)
library(ggplot2, quietly=TRUE)

# Generate an ROC Curve for the rf model on MyDataset [test].
 
MYpr <- predict(MYrf, newdata=na.omit(MYtest[c(MYinput, MYtarget)]), type="prob")[,2]

# Remove observations with missing target.

no.miss  <- na.omit(na.omit(MYtest[c(MYinput, MYtarget)])$STATUS)
miss.list <- attr(no.miss, "na.action")
attributes(no.miss) <- Brown <- NULL

if (length(miss.list)) { pred <- prediction(MYpr[-miss.list], no.miss)} else {pred <- prediction(MYpr, no.miss)}

pe <- performance(pred, "tpr", "fpr")
au <- performance(pred, "auc")@y.values[[1]]
pd <- data.frame(fpr=unlist(pe@x.values), tpr=unlist(pe@y.values))

ggplot(pd, aes(x=fpr, y=tpr)) + geom_line(colour="red") +labs(x= "False Positive Rate", y = "True Positive Rate") +ggtitle("ROC Curve Random Forest MyDataset [test] STATUS") +geom_line(data=data.frame(), aes(x=c(0,1), y=c(0,1)), colour="grey") + annotate("text", x=0.50, y=0.00, hjust=0, vjust=0, size=5,label=paste("AUC =", round(au, 2)))
```

#### Linear Model

```{r}
# Remove observations with missing target.

no.miss   <- na.omit(na.omit(MYtest[c(MYinput, MYtarget)])$STATUS)
miss.list <- attr(no.miss, "na.action")
attributes(no.miss) <- NULL

if (length(miss.list)){pred <- prediction(MYpr[-miss.list], no.miss)} else{ pred <- prediction(MYpr, no.miss)}

performance(pred, "auc")
library(ROCR)
library(ggplot2, quietly=TRUE)

# Generate an ROC Curve for the glm model on MyDataset [test].
 MYpr <- predict(MYglm, type="response", newdata=MYtest[c(MYinput, MYtarget)])
 
 # Remove observations with missing target.

no.miss  <- na.omit(MYtest[c(MYinput, MYtarget)]$STATUS)
miss.list <- attr(no.miss, "na.action")
attributes(no.miss) <- Brown <- NULL
 
if (length(miss.list)){pred <- prediction(MYpr[-miss.list], no.miss)} else{ pred <- prediction(MYpr, no.miss)}

pe <- performance(pred, "tpr", "fpr")
au <- performance(pred, "auc")@y.values[[1]]
pd <- data.frame(fpr=unlist(pe@x.values), tpr=unlist(pe@y.values))

ggplot(pd, aes(x=fpr, y=tpr)) + geom_line(colour="red") +labs(x= "False Positive Rate", y = "True Positive Rate") +ggtitle("ROC Curve Linear Model MyDataset [test] STATUS") +geom_line(data=data.frame(), aes(x=c(0,1), y=c(0,1)), colour="grey") + annotate("text", x=0.50, y=0.00, hjust=0, vjust=0, size=5,label=paste("AUC =", round(au, 2)))
```

### Apply the model 
```{r}
# 2015 Dataset
Employees2016 <- MYtest
ActiveEmployees2016<-subset(Employees2016,STATUS=='ACTIVE')
ActiveEmployees2016$age<-ActiveEmployees2016$age+1
ActiveEmployees2016$length_of_service<-ActiveEmployees2016$length_of_service+1
ActiveEmployees2016$PredictedSTATUS2016<-predict(MYada,ActiveEmployees2016)
 PredictedTerminatedEmployees2016<-subset(ActiveEmployees2016,PredictedSTATUS2016=='TERMINATED')
```

### Total Terminated Employees 
```{r echo= FALSE}

nrow(PredictedTerminatedEmployees2016)
```

### Top 5 records 
```{r echo= FALSE}
head(PredictedTerminatedEmployees2016)
```

#### Export the Dataset
```{r}

write.csv(PredictedTerminatedEmployees2016,file.path(data_dir,"Terminated.csv"))

```

